---
title: "Final Project"
author: "Neomi Rabaev, Yael Snear, Yarin Swisa, Liat Vaizman"
date: "2024-07-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## load the data
```{r}
data <- read_tsv('cleaned_data.tsv')
glimpse(data)
```
### Let's start with visualizing the data

```{r}
ggplot(data, aes(x = langs, y = agg_sci_qual_norm, color = langs)) +
  geom_jitter(width = 0.2, height = 0, size = 2, alpha = 0.6) +
  theme_minimal() +
  labs(
    title = "Normalized Science Scores by Language",
    x = "Language",
    y = "Normalized Science Score"
  ) +  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```


```{r}

# Summarize data to get mean and standard deviation
summary_data <- data %>%
  group_by(langs) %>%
  summarise(
    mean_score = mean(agg_sci_qual_norm),
    sd_score = sd(agg_sci_qual_norm)
  )

# Plotting the data
ggplot(summary_data, aes(x = langs, y = mean_score, fill = langs)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.7) +
  geom_errorbar(aes(ymin = mean_score - sd_score, ymax = mean_score + sd_score), width = 0.2, position = position_dodge(0.7)) +
  theme_minimal() +
  labs(
    title = "Mean Normalized Science Scores by Language",
    x = "Language",
    y = "Mean Normalized Science Score"
  )  +  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

```{r}
ggplot(data, aes(x = langs, y = agg_sci_qual_norm, fill = langs)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    title = "Distribution of Scientific Content Quality by Language",
    x = "Language",
    y = "Quality Score"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```
We can see some surprising result for some languages - German, French and Japanese are lower than we thought, while Turkish and Swahili are higer. Let's look at their scores for each field:
```{r}
# Load necessary packages
library(tidyverse)
library(viridis)

# Filter the data for the specified languages
filtered_data <- data %>%
  filter(langs %in% c("French", "German", "Japanese", "Swahili", "Turkish"))

# Reshape the data to long format
data_long <- filtered_data %>%
  pivot_longer(cols = c(access, contextual, accuracy, how_recent_is_the_information, access_sci),
               names_to = "score_type",
               values_to = "score_value")

# Create the facet graph
ggplot(data_long, aes(x = score_type, y = score_value, fill = score_type)) +
  geom_boxplot() +
  facet_wrap(~ langs) +
  theme_minimal(base_size = 14) +
  labs(title = "Scores by Language",
       x = "Score Type",
       y = "Score Value") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    strip.text = element_text(size = 12, face = "bold"),
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.spacing = unit(2, "lines") # Increase space between facets
  ) +
  scale_fill_brewer(palette = "Set3") +
  theme(plot.margin = margin(20, 20, 20, 20)) 

```
### Divide to train and test
```{r}
library(caret)

# Set the seed for reproducibility
set.seed(123)

# Split the data into 70% training and 30% test
trainIndex <- createDataPartition(data$agg_sci_qual_norm, p = 0.7, 
                                  list = FALSE, 
                                  times = 1)

# Create the training and test data-sets
trainData <- data[trainIndex, ]
testData <- data[-trainIndex, ]
```


```{r}
# Convert langs and country to factors
trainData <- trainData %>%
  mutate(
    langs = factor(langs),
    country = factor(country)
  )

# Run the linear regression model again
model <- lm(agg_sci_qual_norm ~ ., data = trainData)
summary(model)
```
### remove some features from the model
```{r}
model <- lm(agg_sci_qual_norm ~ . - jargon_score - site_type_fct - country - langs, data = trainData)
summary(model)
```


```{r}
model <- lm(agg_sci_qual_norm ~ result_number + link_type + how_recent_is_the_information  + are_there_local_examples_in_the_content_e_g_r + local_examples_prop + does_the_content_present_advantages_and_disad + link_accessibility + jargon_score, data = trainData)
summary(model)
```

```{r}
model1 <- lm(agg_sci_qual_norm ~ . - access_multimedia - access_sci - access - cred - contextual - accuracy,data = trainData)
step_model <- step(model1)
summary(step_model)
```

```{r}
#experimenting more
model2 <- lm(formula = agg_sci_qual_norm ~ langs + type + result_number + 
    link_type + link_accessibility + how_recent_is_the_information + 
    jargon_score + local_examples_prop + does_the_content_present_advantages_and_disad + 
    site_type_fct, data = trainData)
summary(model2)
```

```{r}
predictions <- predict(model2, newdata = testData)
# calculate RMSE
rmse <- function(actual, predicted) {
  sqrt(mean((actual - predicted)^2))
}

# Calculate RMSE for full model
rmse <- rmse(testData$agg_sci_qual_norm, predictions)
cat("RMSE: ", rmse, "\n")

#calculate the MAE
mae <- function(actual, predicted) {
  mean(abs(actual - predicted))
}
mae <- mae(testData$agg_sci_qual_norm, predictions)
cat("MAE: ", mae, "\n")

#calculate the r squared
r_squared <- function(actual, predicted) {
  1 - sum((actual - predicted)^2) / sum((actual - mean(actual))^2)
}
r <- r_squared(testData$agg_sci_qual_norm, predictions)

# Function to calculate adjusted R-squared
adjusted_r_squared <- function(actual, predicted, num_predictors) {
  n <- length(actual)
  r_squared <- 1 - sum((actual - predicted)^2) / sum((actual - mean(actual))^2)
  adj_r_squared <- 1 - (1 - r_squared) * ((n - 1) / (n - num_predictors - 1))
  return(adj_r_squared)
}

cat("r^2: ", r, "\n")
adj_r <- adjusted_r_squared(testData$agg_sci_qual_norm, predictions, 10)

# Print adjusted R-squared
cat("adjusted r^2: ", adj_r)



```

### now, we are trying to remove the rows where link_accessibility=4 (irrelevent)
we want to know if it has an effect on our results


```{r}
data_no_4 <- data %>%
  filter(link_accessibility != 4)
glimpse(data_no_4)
```

```{r}
ggplot(data_no_4, aes(x = langs, y = agg_sci_qual_norm, fill = langs)) +
  geom_boxplot() +
  theme_minimal() +
  labs(
    title = "Distribution of Scientific Content Quality by Language (Withhout Irrelevant Search Results)",
    x = "Language",
    y = "Quality Score"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

### model
```{r}
library(caret)

# Set the seed for reproducibility
set.seed(321)

# Split the data into 70% training and 30% test
trainIndex <- createDataPartition(data_no_4$agg_sci_qual_norm, p = 0.7, 
                                  list = FALSE, 
                                  times = 1)

# Create the training and test data-sets
trainData4 <- data[trainIndex, ]
testData4 <- data[-trainIndex, ]
```

```{r}
model_no_4 <- lm(formula = agg_sci_qual_norm ~ langs + type + result_number + 
    link_type + link_accessibility + how_recent_is_the_information + 
    jargon_score + local_examples_prop + does_the_content_present_advantages_and_disad + 
    site_type_fct, data = trainData4)
summary(model_no_4)

```

```{r}
predictions4 <- predict(model_no_4, newdata = testData4)
# calculate RMSE
rmse <- function(actual, predicted) {
  sqrt(mean((actual - predicted)^2))
}

# Calculate RMSE for full model
rmse <- rmse(testData4$agg_sci_qual_norm, predictions4)
cat("RMSE: ", rmse, "\n")

#calculate the MAE
mae <- function(actual, predicted) {
  mean(abs(actual - predicted))
}
mae <- mae(testData4$agg_sci_qual_norm, predictions4)
cat("MAE: ", mae, "\n")

#calculate the r squared
r_squared <- function(actual, predicted) {
  1 - sum((actual - predicted)^2) / sum((actual - mean(actual))^2)
}
r <- r_squared(testData4$agg_sci_qual_norm, predictions4)

# Function to calculate adjusted R-squared
adjusted_r_squared <- function(actual, predicted, num_predictors) {
  n <- length(actual)
  r_squared <- 1 - sum((actual - predicted)^2) / sum((actual - mean(actual))^2)
  adj_r_squared <- 1 - (1 - r_squared) * ((n - 1) / (n - num_predictors - 1))
  return(adj_r_squared)
}

cat("r^2: ", r, "\n")
adj_r <- adjusted_r_squared(testData4$agg_sci_qual_norm, predictions4, 10)

# Print adjusted R-squared
cat("adjusted r^2: ", adj_r)


```